### 2020-06-29 paper.tex
Updates from Overleaf

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}

\usepackage{graphicx}
\usepackage{ae,aecompl}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{stmaryrd}
\usepackage{amstext}
\usepackage{bm} 
\let\proof\relax
\let\endproof\relax
\usepackage{amsthm}
\usepackage{siunitx}
\usepackage{mathrsfs}
\usepackage{wrapfig}
\usepackage{minted}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{semantic}
\usepackage[dvipsnames]{xcolor}
\usepackage{paralist}
\usepackage{cite}
\usepackage{url}
\usepackage{todonotes}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}
\input{commands}

\begin{document}
%
\title{An FMI-Based Initialization Plugin for INTO-CPS Maestro 2}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Simon Thrane Hansen\inst{1} \and
Casper Thule\inst{1} \and
Claudio Gomes \inst{1}}
%
\authorrunning{S. Thrane et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{DIGIT, Department of Engineering, Aarhus University, \email{\{sth, casper.thule, claudio.gomes\}@eng.au.dk\}}}
%
\maketitle              % typeset the header of the contribution
%

\input{abstract.tex}

\input{introduction.tex}

\input{background.tex}
\section{Calculation of an Initialization Order}\label{sc:initilization}
FMI defines certain information about the initialization order described through different states of a co-simulation. The initialization phase covers the two states (in chronological order):
\begin{itemize}
    \item \textit{Instantiated}
    \item \textit{Initialization Mode}
\end{itemize}
In each of the two states, different groups of FMU variables can be assigned a value. The groups are defined by FMI based on predicates on the characteristics of the variables (\textit{causality}, \textit{initial} and \textit{variability}). These predicates have been extracted and used in the implementation. 
An example of a group is the \textit{INI}-group that consists of FMU variables with $initial = exact\, \lor initial = \,approx $  and $variability \neq constant$. All variables of this group should be set in the \textit{Instantiated}-phase of a co-simulation, and since they have no dependencies/connections to other FMU variables, the order is insignificant. However, all variables of each FMU are group together to perform the fewest possible operations in the initialization.

In the \textit{Initialization Mode} state all variables of all FMUs should be defined.
The variables with \textit{causality = parameter} of each FMU are set first, and the order of which they are set is again insignificant as they are independent.
Afterwards the interconnected variables should be defined, but as stated by the definitions \ref{def:feedthrough}, \ref{def:getout} and \ref{def:setin} the operations \textit{get} and \textit{set} \textbf{require} a specific initialization order, and algebraic loops places even more constraints on the initialization order. This phase should should also support algebraic loops, this support includes utilization of fixed point iteration and detection of non-convergence of algebraic loops.

\subsection{Method to calculate the initialization order}
This section describes the approach to calculate the initialization order of the interconnected FMU variables are based on the the method proposed in Gomes et al. \cite{Gomes2019b, BromanCompositionCo-Simulation}, but the approach used here has been extended with the ability to handle algebraic loops. The approach is formulated in pseudo-code in algorithm \ref{alg:initialization}.
\begin{figure}[htb]
  \centering
    \begin{algorithm}[H]
    \caption{Initialization strategy for Interconnected variables}
    \label{alg:initialization}
      \begin{algorithmic}[1]
        \State $InitializationGraph \gets createGraph(connections)$
        \State $SCCS \gets Tarjan(InitializationGraph)$
        \State $OptimizeInitializationOrder(SCCS)$
        \ForEach {$SCC \in SCCS$}
            \If {$isAlgebraicLoop(SCC)$}
                \State $applyFixedPointIteration(SCC)$;
            \Else
                \State $initializeVariable(SCC)$;
            \EndIf
        \EndFor
        \end{algorithmic}
    \end{algorithm}
\end{figure}

\subsubsection{Structure of the graph}
The initialization algorithm starts by building a directed graph of the dependency between the interconnected variables of the FMUs. The graph is constructed based on the interconnected variables and internal connections (feed-through). Each interconnected variable in the system represents a node, and the edges are based on these connections. The edges of this graph represent precedence constraints based on the algebraic dependencies of the interconnected variables. Please see definition \ref{def:initialization_graph} for a formal definition of the graph.

\begin{definition}[Initialization Graph]\label{def:initialization_graph}
  Given a co-simulation scenario $\tuple{\fmus, \coupling}$, and a set of feed-through dependencies $\bigcup_{c \in \fmus} \set{\feedthrough{c}}$, we define the Initialization Graph where each node represents a port $\outputvar{c} \in \outputs{c}$ or $\inputvar{c} \in \inputs{c}$ of some fmu $c \in \fmus$. The edges are created according to the following rules:
  \begin{compactenum}
    \item For each $c \in \fmus$ and $\inputvar{c} \in \inputs{c}$, if $\coupling(\inputvar{c}) = \outputvar{d}$, add an edge $\outputvar{d} -> \inputvar{c}$.
    \item For each $c \in \fmus$ and $(\inputvar{c},\outputvar{c}) \in \feedthrough{c}$, add an edge $\inputvar{c} -> \outputvar{c}$.
  \end{compactenum}
\end{definition}

As described earlier, not all co-simulation scenarios are valid, and these invalid scenarios need to be identified. This is done by monitoring convergence of all algebraic loops. A valid co-simulation scenario must convergence by definition given in definition \ref{def:convergence}.

\subsubsection{Calculation of an initialization order}
The topological ordering of the graph defined in definition \ref{def:initialization_graph} is used as the initialization order of the interconnected FMU variables. The elements of the topological order are the strongly connected components of the graphs. The strongly connected components are potentially algebraic loops of the system. The algebraic loops are easily identified as the strongly connected components containing more than a single element. The single element members of the topological order are standard interconnected FMU variables, whose port operation should only be performed once during the initialization procedure. \\
The calculation of an initialization order is performed in linear time based on the number of both external and internal connections using Tarjan's algorithm \cite{tarjan_1972}. This algorithm was selected since it solves two of the central issues of the initialization-phase of the co-simulation.
\begin{itemize}
    \item Identifies algebraic loops between interconnected variables (strongly connected components)
    \item Performs a topological sorting of the Strongly Connected Components
\end{itemize}
The algorithm is furthermore well-established, and there exist formal proofs of its correctness and properties\cite{stefan_merz}. The algorithm is among the most efficient graph algorithms for achieving the defined goals.  
Tarjan's algorithm is performing a topological sorting of the strongly connected components of a graph. This makes it able to handle both graphs with and without algebraic loops.

\todo{Come with an example using a graph with loops}

\subsubsection{Handling of algebraic loops}
As described earlier, it is essential to handle the algebraic loops by a particular initialization strategy since they otherwise would invalidate the result of the co-simulation. The strategy for managing algebraic loops is to identify and initialize them using a fixed point iteration until convergence. Because convergence is not guaranteed, it should be monitored, and if convergence can not be established within a finite number of iteration, the co-simulation should be rejected.

\subsection{Optimization of a Initialization Procedure}
An initialization procedure can be optimized by using the fact that the FMI specification allows multiple \textit{set} or \textit{get} operations of the same FMU to be performed in bulk by grouping them together to a single operation over multiple variables with similar characteristic. The criteria of optimization can of an initialization procedure is formalized in definition \ref{def:optimization}
\begin{definition}[Optimization of a Initialization procedure]\label{def:optimization}
  Given an initialization procedure $\sequence{\initcall_i}_{i \in \setnat}$ with a finite ordered sequence of FMU function calls $\functioncall_i \in \allfunctioncalls = \bigcup_{c \in \fmus} \set{\fset{c},\fget{c}},$ and $i$ denoting the order of the function call. It can be optimized if $\exists \functioncall_i, \functioncall_{i +1} \in \allfunctioncalls : \exists c \in \fmus :(\functioncall_i \in {\fset{c}} \land \functioncall_{i+1} \in {\fset{c}}) \lor (\functioncall_i \in {\fget{c}} \land \functioncall_{i+1} \in {\fget{c}})$
\end{definition}
Since an Initialization procedure is defined in the same way as other co-simulation steps, the optimization criteria described in definition \ref{def:optimization} is valid for an arbitrary co-simulation step. \\
The correctness of this optimization can be established by the proof of using the Initialization Graph's topological ordering as the initialization order described by Gomes et al. \cite{Gomes2019}. This proof is still valid since this optimization does not change the structure of the Initialization Graph. \\
A limitation of this optimization strategy is that it is not guaranteed to find all potentially valid optimizations of a co-simulation scenario. Considering it works only on a specific co-simulation step (a topological order of a graph), which is not necessarily unique for a given co-simulation scenario/graph. To perform all viable optimizations of a co-simulation step, a more advanced optimization strategy should be used, or this optimization strategy should be performed on the set of all valid co-simulation steps - yielding a potential very inefficient initialization algorithm.

%This paper's main contribution is to calculate an initialization order of a co-simulation scenario potentially containing algebraic loops. The approach make the circular dependencies between the interconnected converge in the initialization, so the initial values of the interconnected variables in the system being simulated is stable at the time the co-simulation is started.

\subsection{Case study}
Claudio writes this section

\input{realization.tex}

\input{relatedwork.tex}

\input{conclusion.tex}
%
% ---- Bibliography ---

\bibliographystyle{splncs04}
\bibliography{gen_bib}

\end{document}