### 2020-06-29 paper.tex
Added Initialization order file

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}

\usepackage{graphicx}
\usepackage{ae,aecompl}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{stmaryrd}
\usepackage{amstext}
\usepackage{bm} 
\let\proof\relax
\let\endproof\relax
\usepackage{amsthm}
\usepackage{siunitx}
\usepackage{mathrsfs}
\usepackage{wrapfig}
%\usepackage{minted}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{semantic}
\usepackage[dvipsnames]{xcolor}
\usepackage{paralist}
\usepackage{cite}
\usepackage{url}
\usepackage{todonotes}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}
\input{commands}

\begin{document}
%
\title{An FMI-Based Initialization Plugin for INTO-CPS Maestro 2}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Simon Thrane Hansen\inst{1} \and
Casper Thule\inst{1} \and
Claudio Gomes \inst{1}}
%
\authorrunning{S. Thrane et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{DIGIT, Department of Engineering, Aarhus University, \email{\{sth, casper.thule, claudio.gomes\}@eng.au.dk\}}}
%
\maketitle              % typeset the header of the contribution
%

\input{abstract.tex}

\input{introduction.tex}

\input{background.tex}

\input{initializationorder.tex}

%This paper's main contribution is to calculate an initialization order of a co-simulation scenario potentially containing algebraic loops. The approach make the circular dependencies between the interconnected converge in the initialization, so the initial values of the interconnected variables in the system being simulated is stable at the time the co-simulation is started.

\subsection{Case study}
Claudio writes this section

\input{realization.tex}

\input{relatedwork.tex}

\input{conclusion.tex}
%
% ---- Bibliography ---

\bibliographystyle{splncs04}
\bibliography{gen_bib}

\end{document}