### 2020-07-01 background.tex
Updates from Overleaf

\section{Background}\label{sc:background}
In this section, we provide a formalization of FMI co-simulation and a brief background on INTO-CPS Maestro 2.

\subsection{FMU definitions}
To describe the formalization of FMUs, we adopt the vocabulary from \cite{gomes_lucio_vangheluwe_2019, Gomes2018}. The main definitions of relevance to this paper will be presented, but readers are referred to the original publications for more information. This paper is only concerned with the initialization-phase of a co-simulation, making time of an FMU irrelevant. The formalization from Gomes et al. is extended with new definitions regarding algebraic loops, and convergence of fixed point iteration.
\begin{definition}[FMU]\label{def:fmu}
  An FMU with identifier $c$ is represented by the tuple   
  $$\tuple{\stateset{c}, \inputs{c}, \outputs{c}, \fset{c}, \fget{c}},$$
  where:
  \begin{inparadesc}
    \item $\stateset{c}$ represents the state space;
    \item $\inputs{c}$ and $\outputs{c}$ the set of input and output variables, respectively;
    \item $\fset{c} : \stateset{c} \times \inputs{c} \times \values \to \stateset{c}$ and $\fget{c}: \stateset{c} \times \outputs{c} \to \values$ are functions to set the inputs and get the outputs, respectively (we abstract the set of values that each input/output variable can take as $\values$).
  \end{inparadesc}
\end{definition}

\begin{definition}[Scenario]\label{def:cosim_scenario}
  A scenario is a structure $\tuple{\fmus, \coupling}$ where each identifier $c \in \fmus$ is associated with an FMU, as defined in \cref{def:fmu}, and $\coupling(u)=y$ means that the output $y$ is connected to input $u$.
  Let $\allinputs = \bigcup_{c \in \fmus} \inputs{c}$ and $\alloutputs = \bigcup_{c \in \fmus} \outputs{c}$, then $\coupling : \allinputs \to \alloutputs$.
\end{definition}

The following definitions correspond to the operations that are permitted in the initialization phase of a co-simulation.

\begin{definition}[Step]\label{def:cosim_step}
  Given a scenario $\tuple{\fmus, \coupling}$, a co-simulation step, or just step, is a finite ordered sequence of FMU function calls $\sequence{\functioncall_i}_{i \in \setnat} = \functioncall_0, \functioncall_1, \ldots$ with
  $\functioncall_i \in \allfunctioncalls = \bigcup_{c \in \fmus} \set{\fset{c},\fget{c}} \cup (\bigcup_{l \in loops} \fixedpoint{l}),$
  and $i$ denoting the order of the function call.
  Fixed point is defined in \cref{def:fixedpoint} and loop in \cref{def:}.
\end{definition}

\begin{definition}[Initialization]\label{def:initialization}
  Given a scenario $\tuple{\fmus, \coupling}$, we define the initialization procedure $\sequence{\initcall_i}_{i \in \setnat}$ in the same way as a step, with $\initcall_i \in \allfunctioncalls$.
\end{definition}

\begin{definition}[Feed-through]\label{def:feedthrough}
  The input $\inputvar{c} \in \inputs{c}$ feeds through to output $\outputvar{c} \in \outputs{c}$, that is, $(\inputvar{c},\outputvar{c}) \in \feedthrough{c}$, when there exists $v_1, v_2 \in \values$ and $\state{c} \in \stateset{c}$, such that
  $
  \fget{c} (\fset{c}(\state{c}, \inputvar{c}, v_1), \outputvar{c}) \neq \fget{c} (\fset{c}(\state{c}, \inputvar{c}, v_2), \outputvar{c}).
  $
\end{definition}

\begin{definition}[Output Computation]\label{def:getout}
The $\fget{c}(\dontcare, \outputvar{c})$ represents the calculation of output $\outputvar{c}$ of $c \in \fmus$. Given a co-simulation state, it checks whether all inputs that feed-through to $\outputvar{c}$ are defined.
\end{definition}

\begin{definition}[Input Computation]\label{def:setin}
The $\fset{c}(\dontcare, \inputvar{c}, \inputV)$ represents the setting of input $\inputvar{c}$  of $c \in \fmus$. Given a co-simulation state, it checks whether all outputs connected to $\inputvar{c}$ are defined.
\end{definition}

\begin{definition}[Interconnected variable]
An interconnected variable v of a co-simulation scenario $\tuple{\fmus, \coupling}$ is defined as $v \in \allinputs \cup \alloutputs$, then $\coupling : \allinputs \to \alloutputs$.
The set all all interconnected variables is denoted: $V = \allinputs \cup \alloutputs$,
\end{definition}

A graph of the dependencies of a co-simulation scenario is established from the interconnected variables by \cref{def:initialization_graph}. The graph is the foundation for the calculation of the initialization procedure and is therefore referred to as the Initialization Graph.

\begin{definition}[Initialization Graph]\label{def:initialization_graph}
  Given a co-simulation scenario $\tuple{\fmus, \coupling}$, and a set of feed-through dependencies $\bigcup_{c \in \fmus} \set{\feedthrough{c}}$, we define the Initialization Graph where each node represents a port $\outputvar{c} \in \outputs{c}$ or $\inputvar{c} \in \inputs{c}$ of some fmu $c \in \fmus$. The edges are created according to the following rules:
  \begin{compactenum}
    \item For each $c \in \fmus$ and $\inputvar{c} \in \inputs{c}$, if $\coupling(\inputvar{c}) = \outputvar{d}$, add an edge $\outputvar{d} -> \inputvar{c}$.
    \item For each $c \in \fmus$ and $(\inputvar{c},\outputvar{c}) \in \feedthrough{c}$, add an edge $\inputvar{c} -> \outputvar{c}$.
  \end{compactenum}
\end{definition}

The interconnections of FMU variables can lead to circular dependencies between them, as seen in \cref{fig:fmu_cycle} showing a co-simulation scenario containing an algebraic loop and its Initialization Graph.

\begin{figure}
    \centering
    \begin{minipage}{0.55\textwidth}
        \centering
    \includegraphics[width=1\textwidth]{images/fmu_cycle.pdf}
    \end{minipage}\hfill
    \begin{minipage}{0.35\textwidth}
        \centering
    \includegraphics[width=1\textwidth]{images/SCC.pdf}
    \end{minipage}
    \caption{An FMU co-simulation scenario and its Initialization Graph}
    \label{fig:fmu_cycle}
\end{figure}

The following definitions will formalize the concept of an algebraic loop in a co-simulation scenario, and define the problem these algebraic loops are introducing.

\begin{definition}[Algebraic loops] \label{def:loop}
An algebraic loop is defined as a strongly connected component(SCC( of the graph in \cref{def:initialization_graph}. 
A strongly connected component: $SCC = v \in V: Path(v,v)$ and satisfying: $\forall p_1, p_2 \in SCC: Path(p_1, p_2) \land Path(p_2, p_1),$\\
$\loops$ denote the set of all distinct SCCs and $\abs$
Path is defined as the transitive closure of the edges of the graph denoted by \cref{def:initialization_graph}.\\
\end{definition}

The algebraic loops introduce specific challenges in the simulation to handle a non-linear system that can lead to unstable simulation behavior. Algebraic loops are handled using fixed point iterations\cite{Gomes2018} to reach a valid initial state as described in \cref{sec:case_study}. Fixed point iteration is a technique to repeatedly perform the functions of an order sub-sequence of the co-simulation step. The number of repetitions the operations should be performed depends on the scenario. However, the repetitions of the operations are performed until the convergence of the system.

\begin{definition}[Fixed point iteration of an Initialization procedure]\label{def:fixedpoint}
A fixed point iteration of given loop $l$ consist of the order sequence of functions calls:
$\forall v \in l: \exists op \in \fixedpoint{l} : op = \fget{}(\dontcare, v) \lor op = \fset{}(\dontcare, v, \dontcare)$.
Performing a iteration on a co-simulation state leads to a new state to $\fixedpoint{l}(\state{n}) = \state{n + 1}$
The fixed point iteration is performed in order to find a convergent initialization state as defined in \cref{def:convergence}.
\end{definition}

However, a fixed point iteration technique is not guaranteed to convergence if the system is unstable. It means that an upper bound of the number of repetitions needs to be established to ensure termination. In the case of a non-converging algebraic loop, the simulation should be stopped since the co-simulation result would not be trustworthy. The criteria of a valid co-simulation scenario are specified in \cref{def:convergence}.

\begin{definition}[Convergence of Fixed point iteration]\label{def:convergence}
A fixed point iteration converges if a finite number of iterations will make the difference of the output value of the same operation between two following iterations within a certain threshold.\\
$Convergence = \exists n \in \setnat: \forall l \in \loops \implies |f(p, n + 1) - f(p, n)| \leq T$\\
$f$ is the operation performed on a loop-variable p of a given iteration n.
\end{definition}

\subsection{INTO-CPS Maestro 2}\todo{We can have some more in this section - Ask Casper if he can do it?}
INTO-CPS Maestro 2\footnote{currently in alpha \url{https://github.com/INTO-CPS-Association/maestro/tree/2.0.0-alpha}}\cite{thule_maestro2_2019} is an FMI-based co-simulation framework set to supersede Maestro\cite{Maestro}. The philosophy of the framework is to apply plugins to generate co-simulation specifications expressed in the domain specific language called Maestro Base Language (MaBL). Such specifications are then interpreted and executed, resolving in the execution of a co-simulation.