### 2020-11-13 initializationorder.tex

\section{Calculation of an Initialization Order}\label{sc:initilization}
The FMI specification defines certain information about the initialization order described through different states of a co-simulation. The initialization phase covers the two states (in chronological order) defined in the FMI specification:
\begin{itemize}
    \item \textit{Instantiated}
    \item \textit{Initialization Mode}
\end{itemize}
In each of the two states, different groups of FMU variables and parameters are potentially assigned a value. The groups are defined by FMI based on the characteristics of the FMU variables. The rules have been extracted as predicates and used in the implementation. 
Some groups consist of variables and parameters whose value does not depend on other variables. These independent variables and parameters can be set in the \textit{Instantiated} phase of the Initialization. Since these variables have no connections to other FMU variables - meaning they are not represented in the graph of \cref{def:initialization_graph}, the order their value is set in is insignificant. 
The setting and getting operations of each FMU are grouped to perform the fewest possible FMU-operations during the Initialization.

In the \textit{Initialization Mode} state all the interconnected variable is being defined, but as stated by the \cref{def:feedthrough,def:getout,def:setin} the operations \textit{get} and \textit{set} \textbf{require} that the operations are performed in a specific order. Furthermore, algebraic loops place even more requirements on the initialization strategy. Since each non-trivial strongly connected component (algebraic loop) needs to be isolated from the other variables of the system to calculate their initial values using fixed point iteration as described in \cref{def:fixedpoint}. 
After the \textit{Initialization Mode} state, all variables of all FMUs in the co-simulation scenario should be defined, and the co-simulation should be ready to enter the \textit{modelInitialized} state.

%\%claudio{I feel like the previous two sentences are a bit contradictory, and are probably constructed like so because you did not deal with algebraic loops upfront. When algebraic loops are not there, there must be an order. But when algebraic loops are there, it's not that there's more constraints... it's just that one has to isolate the strong components from the rest. The result is then an order between strongly connected components, but no order within each component}

\subsection{Method to Calculate the Initialization Order}
This section describes the approach taken to calculate the initialization order of the interconnected FMU variables. The approach is based on the strategy proposed in Gomes et al. \cite{Gomes2019b, BromanCompositionCo-Simulation}, but the approach in this work is extended with the ability to handle the Initialization of algebraic loops.

%\claudio{I like the way in which you've structured this subsection... but perhaps you don't need to make this structure explciit in the latex.. just leave the subsubsection comments commented out, but still follow the structure.}
%\subsubsection{Structure of the graph}
The initialization algorithm starts by building a directed graph of the dependencies between the interconnected variables of the FMUs. The graph is constructed based on the interconnected variables and internal connections (feed-through); please see \cref{def:initialization_graph} for a formal definition of the graph. 
%Each interconnected variable in the system represents a node, and the edges are based on these connections. The edges of this graph represent precedence constraints based on the algebraic dependencies of the interconnected variables. Please see \cref{def:initialization_graph} for a formal definition of the graph.

%As described earlier, not all co-simulation scenarios are suitable, and these invalid scenarios need to be identified in this phase to avoid wasting valuable development. This is accomplished by monitoring convergence of all algebraic loops. A valid co-simulation scenario must convergence by \cref{def:convergence}.
%\claudio{Wait a second... you can only monitor convergence once you have an initialization order, and you start solving the loops. Why is paragraph here if the calculation of an initialization order is still to be explained? Maybe move this explanation to after that part.}

%\subsubsection{Calculation of an initialization order}
The topological ordering of the strongly connected components of the graph defined in \cref{def:initialization_graph} is the initialization order of the interconnected FMU variables. 
The non-trivial strongly connected components are algebraic loops of the system. The trivial ones are standard interconnected FMU variables, whose port operation should be performed only once during the initialization procedure.
The calculation of an initialization order is performed in linear time based on the number of external and internal connections using Tarjan's algorithm \cite{tarjan_1972}. 
%This algorithm is selected due to its properties. It solves two of the central issues of the initialization-phase of the co-simulation.
%\begin{itemize}
%    \item Identifies algebraic loops between interconnected variables (strongly connected components)
%    \item Performs a topological sorting of the Strongly Connected Components
%\end{itemize}
%The algorithm is well-established, and there exist formal proofs of its correctness and properties\cite{stefanMerz}. 
% The algorithm is among the most efficient graph algorithms for accomplishing the defined goals.
% Tarjan's algorithm is performing a topological sorting of the strongly connected components of a graph. Moreover, it can handle both graphs with and without algebraic loops.

%\subsubsection{Handling of algebraic loops}
As described in earlier sections, it is essential to handle algebraic loops by a particular initialization strategy since the loops otherwise would invalidate the co-simulation result. The procedure for initializing algebraic loops is identifying and initializing them using a fixed point iteration strategy until convergence. Since convergence is not guaranteed, this property is monitored using \cref{def:convergence} to see if the difference between all the output variables of two successive iterations is below a defined threshold. Suppose convergence is not established within a finite number of iterations\footnote{5 iterations is the default in our approach. This number is based on experience.}, the co-simulation scenario is rejected to avoid running an invalid simulation.

\subsection{Optimization of a Initialization Procedure}
An initialization procedure can sometimes be optimized since the FMI specification allows multiple \textit{set} or \textit{get} operations of the same FMU to be performed in bulk by grouping them together to a single operation over multiple variables with similar characteristic. This criteria of optimization is formalized in \cref{def:optimization}.
\begin{definition}[Optimization of a Initialization procedure]\label{def:optimization}
  Given an initialization procedure $\sequence{\initcall_i}_{i \in \setnat}$ with a finite ordered sequence of FMU function calls $\functioncall_i \in \allfunctioncalls = \bigcup_{c \in \fmus} \set{\fset{c},\fget{c}},$ and $i$ denoting the order of the function call. It can be optimized if $\exists \functioncall_i, \functioncall_{i +1} \in \allfunctioncalls : \exists c \in \fmus :(\functioncall_i \in {\fset{c}} \land \functioncall_{i+1} \in {\fset{c}}) \lor (\functioncall_i \in {\fget{c}} \land \functioncall_{i+1} \in {\fget{c}})$
\end{definition}
%Since an Initialization procedure is defined in the same way as other co-simulation steps (see \cref{def:initialization}), the optimization criteria described in \cref{def:optimization} is valid for an arbitrary co-simulation step. \\
The correctness of the optimization in \cref{def:optimization} is established by the proof of using the Initialization Graph's topological ordering as the initialization order by Gomes et al. \cite{Gomes2019}. Their proof is trivially shown to cover this approach since the optimization does not change the structure of the Initialization Graph. \\
A limitation of this optimization strategy is that it is not guaranteed to find all potentially valid optimizations of a co-simulation scenario. Considering it works only on a specific co-simulation step (a topological order of a graph), which is not necessarily unique for a given co-simulation scenario. A more advanced optimization strategy needs to be developed to perform all viable optimizations of a co-simulation step. Another solution is to apply this optimization strategy on the set of all valid co-simulation steps - yielding a potential very inefficient initialization algorithm.
The initialization of a co-simulation is typically not the most time consuming or computational heaviest part of the co-simulation. However, it is still considered a low hanging fruit to apply this optimization to optimize the initialization-

\subsection{The Complete Initialization Strategy}
The pseudo-code in \cref{alg:initialization} formulates the entire initialization strategy of the interconnected variables of a co-simulation scenario.
\begin{figure}[H]
  \centering
    \begin{algorithm}[H]
    \caption{Initialization strategy for Interconnected variables}
    \label{alg:initialization}
      \begin{algorithmic}[1]
        \State $InitializationGraph \gets createGraph(connections)$
        \State $SCCS \gets Tarjan(InitializationGraph)$
        \State $OptimizeInitializationOrder(SCCS)$
        \ForEach {$SCC \in SCCS$}
            \If {$isAlgebraicLoop(SCC)$}
                \State $applyFixedPointIteration(SCC)$;
            \Else
                \State $initializeVariable(SCC)$;
            \EndIf
        \EndFor
        \end{algorithmic}
    \end{algorithm}
\end{figure}

As seen from the algorithm in \cref{alg:initialization}, the algebraic loops are handled using a different initialization strategy compared to the other trivial SCC of a single interconnected FMU variable.